Generating Random Logic Programs Using Constraint Programming

### 2020-01-29 paper/paper.tex
Finishing touches

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

@@

\subsection{PRELIMINARIES}

@@

\subsection{PARAMETERS}

@@

We also define a measure of how complicated a body of a clause can become. As
each body is represented by a tree (see \cref{sec:bodies}), we set
$\maxNumNodes{} \ge 1$ to be the maximum number of nodes in the tree
representation of any clause. We also set $\maxNumClauses{}$ to be the maximum
number of clauses in a program. We must have that $\maxNumClauses{} \ge
|\predicates{}|$ because we require each predicate to have at least one clause
that defines it. The model supports eliminating either all cycles or just
negative cycles (see \cref{sec:cycles}) and enforcing predicate independence
(see \cref{sec:independence}), so a list of independent pairs of predicates is
another parameter. Since this model can generate probabilistic as well as
non-probabilistic programs, each clause is paired with a probability which is
randomly selected from a given list (which is our last parameter). For
generating non-probabilistic programs, one can set this list equal to $\{ 1 \}$.
Finally, we define $\tokens{} = \{ \neg, \land, \lor, \top \}$ as the set of
tokens that (together with atoms) form a clause. All decision variables of the
model can now be divided into $2 \times \maxNumClauses{}$ separate groups,
treating the body and the head of each clause separately. We say that the
variables are contained in two arrays: $\arrayt{bodies}{\maxNumClauses{}}{Body}$
and $\arrayt{heads}{\maxNumClauses{}}{Head}$. Since the order of the clauses
does not change the meaning of the program, we can also state our first
constraint:

@@

Here, $\top$ is the value we use for the remaining one-node trees. The
elements of the $\variable{values}$ array are \emph{nodes}.

@@

Furthermore, we can use \cref{constr:arity} again to disable the extra
arguments.

@@

Given any clause, we can permute the variables in it without changing the
meaning of the clause or the entire program. Thus, we want to fix the order of
variables to eliminate unnecessary symmetries. Informally, we can say that
variable $X$ goes before variable $Y$ if its first occurrence in either the head
or the body of the clause is before the first occurrence of $Y$. Note that the
constrains described in this section only make sense if $|\variables| > 1$. Also
note that all definitions and constraints here are on a per-clause basis.

@@

  Then we can use the $\variable{setsIntsChanneling}$ constraint
  to define $\variable{occ}[|\constants{}| + |\variables{}| + 1]$ as an
  array of subsets of $\{ 0, \dots, N-1 \}$ such that for all $i = 0, \dots, N
  - 1$, and $t \in \constants{} \cup \variables{} \cup \{ \Box \}$,
  \[
    i \in \variable{occ}[t] \quad \iff \quad
    \variable{terms}[i] = t
  \]
\end{definition}

@@

Here, a value of zero means that the variable does not occur in the clause. We
want to use specifically zero for this so that we could use
\cref{constraint:diffbutzero} later. Because of this choice, the definition of
$\variable{intros}$ shifts all indices by one. Lastly, we add the constraint
that eliminates variable symmetries:

@@

The reason why we used zero to represent a variable not being introduced is so
that we could efficiently rephrase \cref{constraint:all_diff} for the
$\variable{intros}$ array:

@@

Using this new variable, we can easily add a constraint saying that
non-predicate nodes in the tree representation of a clause cannot have variables
as arguments:

@@

\begin{example} \label{ex:program}
  Consider the following (fragment of a) program:
  \begin{align}
    \mathsf{sibling}(X, Y) &\gets \mathsf{parent}(X, Z) \land \mathsf{parent}(Y, Z), \nonumber \\
    \mathsf{father}(X, Y) &\gets \mathsf{parent}(X, Y) \land \neg\mathsf{mother}(X, Y). \label[clause]{eq:example_clause}
  \end{align}
  Its predicate dependency graph is in \cref{fig:predicate_dependencies}.
  Because of the negation in \cref{eq:example_clause} (as seen in
  \cref{fig:example_tree2}), the edge from $\mathsf{mother}$ to
  $\mathsf{father}$ is negative, while the other two edges are positive.

@@

We can now add a constraint to define an adjacency matrix for the predicate
dependency graph but without positivity/negativity.

@@

Each entailment algorithm returns one out of three different values:
\textsc{true} if the constraint is guaranteed to hold, \textsc{false} if the
constraint is violated, and \textsc{undefined} if whether the constraint will be
satisfied or not depends on the future decisions made by the solver.
\cref{alg:independence_entailment} outlines a simple entailment algorithm for
the independence of two predicates $p_1$ and $p_2$. First, we separately
calculate all dependencies of $p_1$ and $p_2$, and look at the set $D$ of
dependencies that $p_1$ and $p_2$ have in common. If there are none, then the
predicates are clearly independent. If they have a dependency in common that is
already fully determined ($\Determined$) for both predicates, then they cannot be
independent. Otherwise, we return \textsc{undefined}.

@@

Propagation algorithms have two goals: causing a contradiction (failing) in
situations where the corresponding entailment algorithm would return
\textsc{false}, and eliminating values from domains of variables that are
guaranteed to cause a contradiction. \cref{alg:independence_propagation} does
the former on \cref{line:fail}. Furthermore, for any dependency shared between
predicates $p_1$ and $p_2$, if it is determined ($\Determined$) for one
predicate and almost determined ($\AlmostDetermined$) for another, then the edge
that prevents the $\AlmostDetermined$ from becoming a $\Determined$ cannot
exist--\cref{line:propagation_condition,line:propagation_result} take care of
that.

@@

The function $\getDependencies$ in \cref{alg:dependencies} calculates $D_p$ for
any predicate $p$. It has two versions: $\getDependencies(p, 1)$ returns all
dependencies, while $\getDependencies(p, 0)$ returns only determined and
almost-determined dependencies. It starts by establishing the predicate $p$
itself as a dependence and continues to add dependencies of dependencies until
the set $D$ stabilises. For each dependency $d \in D$, we look at the in-links
of $d$ in the predicate dependency graph. If the edge from some predicate $q$ to
$d.\mathsf{predicate}$ is fully determined and $d$ is determined, then $q$ is
another determined dependency of $p$. If the edge is determined but $d$ is
almost determined, then $q$ is an almost-determined dependency. The same outcome
applies if $d$ is fully determined but the edge is undetermined. Finally, if we
are interested in collecting all dependencies regardless of their status, then
$q$ is a dependency of $p$ as long as the edge from $q$ to
$d.\mathsf{predicate}$ is possible. Note that if there are multiple paths in the
dependency graph from $q$ to $p$, \cref{alg:dependencies} could include $q$ once
for each possible type ($\Determined$, $\Undetermined$, and
$\AlmostDetermined$), but
\cref{alg:independence_propagation,alg:independence_entailment} would still work
as intended.

@@

\begin{example} \label{example:independence}
  Consider this partially determined (fragment of a) program:
  \begin{align*}
    \Box(X, Y) &\gets \mathsf{parent}(X, Z) \land \mathsf{parent}(Y, Z),\\
    \mathsf{father}(X, Y) &\gets \mathsf{parent}(X, Y) \land \neg\mathsf{mother}(X, Y)
  \end{align*}
  where $\Box$ indicates an unknown predicate with domain
  \[
    D = \{ \mathsf{father}, \mathsf{mother}, \mathsf{parent}, \mathsf{sibling}
    \}.
  \]
  The predicate dependency graph without positivity/negativity (as defined in
  \cref{def:adjacency_matrix}) is represented in
  \cref{fig:dependencies_matrix,fig:dependencies2}.

  Suppose we have a constraint that $\mathsf{mother}$ and $\mathsf{parent}$ must
  be independent. The lists of potential dependencies for both predicates are:
  \begin{align*}
    D_{\mathsf{mother}} &= \{ \Determined(\mathsf{mother}), \AlmostDetermined(\mathsf{parent}, \mathsf{parent}, \mathsf{mother}) \}, \\
    D_{\mathsf{parent}} &= \{ \Determined(\mathsf{parent}) \}.
  \end{align*}
  An entailment check at this stage would produce \textsc{undefined}, but
  propagation replaces the boxed value in \cref{fig:dependencies_matrix} with
  zero, eliminating the potential edge from $\mathsf{parent}$ to
  $\mathsf{mother}$. This also eliminates $\mathsf{mother}$ from $D$, and,
  although some undetermined variables remain, this is enough to make
  \cref{alg:independence_entailment} return \textsc{true}.
\end{example}

@@

Having no negative cycles in the predicate dependency graph is a requirement for
probabilistic logic programming language ProbLog \citep{kimmig2009trading},
although it has been shown how the requirement can be alleviated by introducing
negative probabilities \citep{DBLP:journals/ijar/BuchmanP17}. Ideally, we would
like to design a constraint for that similar to the constraint for independence
in the previous section. However, the difficulty with creating a propagation
algorithm for negative cycles is that there seems to be no good way to extend
\cref{def:adjacency_matrix} so that the adjacency matrix captures
positivity/negativity. Thus, we settle for an entailment algorithm with no
propagation.

@@

If the algorithm finds a negative cycle in this fully-determined part of the
program, then it is unavoidable and we must return \textsc{false}. If there was
no negative cycle and the entire program is determined-enough to contribute to
the graph, then there cannot be any negative cycles. In all other cases, it
is too early to tell.

@@

\begin{figure*}
  \centering
  \input{phase_transition.tex}
  \caption{The distribution of the number of nodes in the binary search tree as
    a function of the number of independent pairs of predicates for
    $|\predicates| = 8$. Significant outliers are hidden, the dots denote mean
    values, and the vertical axis is on a $\log_{10}$ scale.}
  \label{fig:phase_transition}
\end{figure*}

@@

We ran close to \num{400000} experiments, investigating whether the model is
efficient enough to generate reasonably-sized programs and gaining insight into
what parameter values make the constraint satisfaction problem harder. For these
experiments, we use Choco~4.10.2 \citep{choco} with Java~8 on a personal
computer with an Intel Core i5-8250U processor and Arch Linux 5.4.12-arch1-1
operating system. For $|\predicates{}|$, $|\variables{}|$, $|\constants{}|$,
$\maxNumNodes{}$, and $\maxNumClauses{} - |\predicates{}|$ (i.e., the number of
clauses in addition to the mandatory $|\predicates{}|$ clauses), we assign all
combinations of 1, 2, 4, 8. $\maxArity{}$ is assigned to values 1--4. For each
$|\predicates{}|$, we also iterate over all possible numbers of independent
pairs of predicates, ranging from 0 up to $\binom{|\predicates{}|}{2}$. For each
combination of the above-mentioned parameters, we pick ten random pairs of:
$\arities{}$ (such that $\maxArity{}$ occurs at least once) and the required
number of independent pairs (without repetition). We then run the solver with a
\SI{60}{\second} timeout.

@@

In \cref{fig:impact}, we plot how the mean number of nodes in the binary search
tree grows as a function of each parameter (the plot for the median is very
similar). The growth of each curve suggest how well/poorly the model scales with
higher values of the parameter. From this plot, it is clear that
$\maxNumNodes{}$ is the limiting factor. This is because some tree structures
can be impossible to fill with predicates without creating either a negative
cycle or a forbidden dependence, and such trees become more common as the number
of nodes increases. Likewise, a higher number of predicates complicates the
situation as well.

@@

% TODO: the positive side (what did we achieve?)
%We were able to design an efficient model

One major question that was not addressed in this paper is that of logical
equivalence. We eliminated permutation symmetries and constrained variable usage
so that the same `functionality' is always assigned to the same variable, but
syntactically different logical formulas are always treated as being different.
For example, all three of these formulas are logically equivalent but
syntactically different: $\neg(\mathsf{P}(X) \lor \mathsf{Q}(X))$,
$\neg\mathsf{P}(X) \land \neg\mathsf{Q}(X)$, and $\neg\mathsf{Q}(X) \land
\neg\mathsf{P}(X)$. While there are many situations where one would prefer to
treat logically equivalent formulas as symmetries that should be eliminated,
there are also situations where we do care about enumerating different ways to
express the same probability distribution or knowledge base. For instance,
one could look for the `best' logically equivalent formula, or investigate
whether some formulas result in faster inference than others.

Another issue worth discussing is that of randomness. Most random models sample
from a well-defined probability distribution whereas a constraint model, while
potentially sufficiently random, provides no such guarantees. In future work, it
would be interesting to examine the probability distribution of logic programs
as generated by our model, although it is not obvious how such a
probability distribution could be characterised. One example of a clear bias
exhibited by our model is that of complexity over simplicity, i.e., if a tree
that represents the body of a clause can have up to $\maxNumNodes{}$
nodes---because the number of possible trees with $n$ nodes grows with $n$---it
is much more likely to have $\maxNumNodes{}$ or $\maxNumNodes{} - 1$ nodes than,
e.g., one or two.

\subsubsection*{Acknowledgements}

% The author would like to thank Vaishak Belle for his comments.
This work was supported by the EPSRC Centre for Doctoral Training in Robotics
and Autonomous Systems, funded by the UK Engineering and Physical Sciences
Research Council (grant EP/S023208/1).

\newpage

>>>>

\author{}

%\author{ {\bf Paulius Dilkas} \\
%School of Informatics \\
%University of Edinburgh \\
%Edinburgh, United Kingdom \\
%}

@@

\section{PRELIMINARIES}

@@

\subsection{PARAMETERS OF THE MODEL}

@@

We also define a measure of how complicated a body of a clause can become. As
each body is represented by a tree (see \cref{sec:bodies}), we set
$\maxNumNodes{} \ge 1$ to be the maximum number of nodes in the tree
representation of any clause. We also set $\maxNumClauses{}$ to be the maximum
number of clauses in a program. We must have that $\maxNumClauses{} \ge
|\predicates{}|$ because we require each predicate to have at least one clause
that defines it. The model supports eliminating either all cycles or just
negative cycles (see \cref{sec:cycles}) and enforcing predicate independence
(see \cref{sec:independence}), so a list of independent pairs of predicates is
another parameter. Since this model can generate probabilistic as well as
non-probabilistic programs, each clause is paired with a probability which is
randomly selected from a given list (i.e., our last parameter). For
generating non-probabilistic programs, one can set this list equal to $\{ 1 \}$.
Finally, we define $\tokens{} = \{ \neg, \land, \lor, \top \}$ as the set of
tokens that (together with atoms) form a clause. All decision variables of the
model can now be divided into $2 \times \maxNumClauses{}$ separate groups,
treating the body and the head of each clause separately. We say that the
variables are contained in two arrays: $\arrayt{bodies}{\maxNumClauses{}}{Body}$
and $\arrayt{heads}{\maxNumClauses{}}{Head}$. Since the order of the clauses
does not change the meaning of the program, we can also state our first
constraint:

@@

Here, $\top$ is the value we use for the remaining one-node trees. The
elements of the $\variable{values}$ array are nodes:

@@

Furthermore, we can use \cref{constr:arity} to again disable the extra
arguments.

@@

Given any clause, we can permute the variables in that clause without changing
the meaning of the clause or the entire program. Thus, we want to fix the order
of variables to eliminate unnecessary symmetries. Informally, we can say that
variable $X$ goes before variable $Y$ if the first occurrence of $X$ in either
the head or the body of the clause is before the first occurrence of $Y$. Note
that the constrains described in this section only make sense if $|\variables| >
1$. Also note that all definitions and constraints here are on a per-clause
basis.

@@

  Then we can use a channeling constraint to define
  $\variable{occ}[|\constants{}| + |\variables{}| + 1]$ as an array of subsets
  of $\{ 0, \dots, N-1 \}$ such that for all $i = 0, \dots, N - 1$, and $t \in
  \constants{} \cup \variables{} \cup \{ \Box \}$,
  \[
    i \in \variable{occ}[t] \quad \iff \quad
    \variable{terms}[i] = t
  \]
\end{definition}

@@

Here, a value of zero means that the variable does not occur in the clause. The
reason why we want to use specifically zero for this will become clear with
\cref{constraint:diffbutzero}. Because of this choice, the definition of
$\variable{intros}$ shifts all indices by one. Lastly, we add the constraint
that eliminates variable symmetries:

@@

The reason why we used zero to represent an unused variable is so that we could
efficiently rephrase \cref{constraint:all_diff} for the $\variable{intros}$
array:

@@

Using this new variable, we can add a constraint saying that non-predicate nodes
in the tree representation of a clause cannot have variables as arguments:

@@

\begin{example} \label{ex:program}
  Consider the following (fragment of a) program:
  \begin{align}
    \mathsf{sibling}(X, Y) &\gets \mathsf{parent}(X, Z) \land \mathsf{parent}(Y, Z), \nonumber \\
    \mathsf{father}(X, Y) &\gets \mathsf{parent}(X, Y) \land \neg\mathsf{mother}(X, Y) \label[clause]{eq:example_clause}
  \end{align}
  Its predicate dependency graph is in \cref{fig:predicate_dependencies}.
  Because of the negation in \cref{eq:example_clause} (as seen in
  \cref{fig:example_tree2}), the edge from $\mathsf{mother}$ to
  $\mathsf{father}$ is negative, while the other two edges are positive.

@@

We can now add a constraint to define an adjacency matrix for the predicate
dependency graph but without positivity/negativity:

@@

Each entailment algorithm returns one out of three different values:
\textsc{true} if the constraint is guaranteed to hold, \textsc{false} if the
constraint is violated, and \textsc{undefined} if whether the constraint will be
satisfied or not depends on the future decisions made by the solver.
\cref{alg:independence_entailment} outlines a simple entailment algorithm for
the independence of two predicates $p_1$ and $p_2$. First, we separately
calculate all dependencies of $p_1$ and $p_2$ and look at the set $D$ of
dependencies that $p_1$ and $p_2$ have in common. If there are none, then the
predicates are clearly independent. If they have a dependency in common that is
already fully determined ($\Determined$) for both predicates, then they cannot be
independent. Otherwise, we return \textsc{undefined}.

@@

Propagation algorithms have two goals: causing a contradiction (failing) in
situations where the corresponding entailment algorithm would return
\textsc{false}, and eliminating values from domains of variables that are
guaranteed to cause a contradiction. \cref{alg:independence_propagation} does
the former on \cref{line:fail}. Furthermore, for any dependency shared between
predicates $p_1$ and $p_2$, if it is determined ($\Determined$) for one
predicate and almost determined ($\AlmostDetermined$) for another, then the edge
that prevents the $\AlmostDetermined$ from becoming a $\Determined$ cannot
exist--\cref{line:propagation_condition,line:propagation_result} handle this
possibility.

@@

The function $\getDependencies$ in \cref{alg:dependencies} calculates $D_p$ for
any predicate $p$. It has two versions: $\getDependencies(p, 1)$ returns all
dependencies, while $\getDependencies(p, 0)$ returns only determined and
almost-determined dependencies. It starts by establishing the predicate $p$
itself as a dependency and continues to add dependencies of dependencies until
the set $D$ stabilises. For each dependency $d \in D$, we look at the in-links
of $d$ in the predicate dependency graph. If the edge from some predicate $q$ to
$d.\mathsf{predicate}$ is fully determined and $d$ is determined, then $q$ is
another determined dependency of $p$. If the edge is determined but $d$ is
almost determined, then $q$ is an almost-determined dependency. The same outcome
applies if $d$ is fully determined but the edge is undetermined. Finally, if we
are interested in collecting all dependencies regardless of their status, then
$q$ is a dependency of $p$ as long as the edge from $q$ to
$d.\mathsf{predicate}$ is possible. Note that if there are multiple paths in the
dependency graph from $q$ to $p$, \cref{alg:dependencies} could include $q$ once
for each possible type ($\Determined$, $\Undetermined$, and
$\AlmostDetermined$), but
\cref{alg:independence_propagation,alg:independence_entailment} would still work
as intended.

@@

\begin{example} \label{example:independence}
  Consider this partially determined (fragment of a) program:
  \begin{align*}
    \Box(X, Y) &\gets \mathsf{parent}(X, Z) \land \mathsf{parent}(Y, Z),\\
    \mathsf{father}(X, Y) &\gets \mathsf{parent}(X, Y) \land \neg\mathsf{mother}(X, Y)
  \end{align*}
  where $\Box$ indicates an unknown predicate with domain
  \[
    D_\Box = \{ \mathsf{father}, \mathsf{mother}, \mathsf{parent},
    \mathsf{sibling} \}.
  \]
  The predicate dependency graph without positivity/negativity (as defined in
  \cref{def:adjacency_matrix}) is represented in
  \cref{fig:dependencies_matrix,fig:dependencies2}.

  Suppose we have a constraint that $\mathsf{mother}$ and $\mathsf{parent}$ must
  be independent. The lists of potential dependencies for both predicates are:
  \begin{align*}
    D_{\mathsf{mother}} &= \{ \Determined(\mathsf{mother}), \AlmostDetermined(\mathsf{parent}, \mathsf{parent}, \mathsf{mother}) \}, \\
    D_{\mathsf{parent}} &= \{ \Determined(\mathsf{parent}) \}.
  \end{align*}
  An entailment check at this stage would produce \textsc{undefined}, but
  propagation replaces the boxed value in \cref{fig:dependencies_matrix} with
  zero, eliminating the potential edge from $\mathsf{parent}$ to
  $\mathsf{mother}$. This also eliminates $\mathsf{mother}$ from $D_\Box$, and,
  although some undetermined variables remain, this is enough to make
  \cref{alg:independence_entailment} return \textsc{true}.
\end{example}

@@

Having no negative cycles in the predicate dependency graph is a requirement for
probabilistic logic programming language ProbLog \citep{kimmig2009trading},
although it has been shown how the requirement can be alleviated by introducing
negative probabilities \citep{DBLP:journals/ijar/BuchmanP17}. Ideally, we would
like to design a constraint for negative cycles similar to the constraint for
independence in the previous section. However, the difficulty with creating a
propagation algorithm for negative cycles is that there seems to be no good way
to extend \cref{def:adjacency_matrix} so that the adjacency matrix captures
positivity/negativity. Thus, we settle for an entailment algorithm with no
propagation.

@@

If the algorithm finds a negative cycle in this fully-determined part of the
program, then we must return \textsc{false}. If there was no negative cycle and
the entire program is (sufficiently) determined, then there cannot be any
negative cycles. In all other cases, it is too early to tell.

@@

\begin{figure*}
  \centering
  \input{phase_transition.tex}
  \caption{The distribution of the number of nodes in the binary search tree as
    a function of the number of independent pairs of predicates for
    $|\predicates| = 8$. Outliers are hidden, the dots denote mean values, and
    the vertical axis is on a $\log_{10}$ scale.}
  \label{fig:phase_transition}
\end{figure*}

@@

We ran close to \num{400000} experiments, investigating whether the model is
efficient enough to generate reasonably-sized programs and gaining insight into
what parameter values make the constraint satisfaction problem harder. For these
experiments, we use Choco~4.10.2 \citep{choco} with Java~8 on a computer with an
Intel Core i5-8250U processor. For $|\predicates{}|$, $|\variables{}|$,
$|\constants{}|$, $\maxNumNodes{}$, and $\maxNumClauses{} - |\predicates{}|$
(i.e., the number of clauses in addition to the mandatory $|\predicates{}|$
clauses), we assign all combinations of 1, 2, 4, 8. $\maxArity{}$ is assigned to
values 1--4. For each $|\predicates{}|$, we also iterate over all possible
numbers of independent pairs of predicates, ranging from 0 up to
$\binom{|\predicates{}|}{2}$. For each combination of the above-mentioned
parameters, we pick ten random ways to assign arities to predicates (such that
$\maxArity{}$ occurs at least once) and ten random combinations of independent
pairs. We then run the solver with a \SI{60}{\second} timeout.

@@

In \cref{fig:impact}, we plot how the mean number of nodes in the binary search
tree grows as a function of each parameter (the plot for the median is very
similar). The growth of each curve suggest how well/poorly the model scales with
higher values of the parameter. From this plot, it is clear that
$\maxNumNodes{}$ is the limiting factor. This is because some tree structures
can be impossible to fill with predicates without creating either a negative
cycle or a forbidden dependency, and such trees become more common as the number
of nodes increases. Likewise, a higher number of predicates complicates the
situation as well.

@@

We were able to design an efficient model for generating both logic programs and
probabilistic logic programs. The model avoids unnecessary symmetries, generates
valid programs, and can ensure predicate independence. Note that there is one
kind of symmetries not handled by our model, i.e., logical equivalence. For
example, all three of these formulas are logically equivalent but syntactically
different: $\neg(\mathsf{P}(X) \lor \mathsf{Q}(X))$, $\neg\mathsf{P}(X) \land
\neg\mathsf{Q}(X)$, and $\neg\mathsf{Q}(X) \land \neg\mathsf{P}(X)$. While there
are many situations where one would prefer to treat logically equivalent
formulas as symmetries that should be eliminated, there are also situations
where we do care about enumerating different ways to express the same
probability distribution or knowledge base. For instance, one could look for the
`best' logically equivalent formula, or investigate whether some formulas result
in faster inference than others.

Another issue worth discussing is that of randomness. Most random models sample
from a well-defined probability distribution whereas a constraint model, while
potentially sufficiently random, provides no such guarantees. In future work, it
would be interesting to examine the probability distribution of logic programs
as generated by our model, although it is not obvious how such a
probability distribution could be characterised.% One example of a clear bias
%exhibited by our model is that of complexity over simplicity, i.e., if a tree
%that represents the body of a clause can have up to $\maxNumNodes{}$
%nodes---because the number of possible trees with $n$ nodes grows with $n$---it
%is much more likely to have $\maxNumNodes{}$ or $\maxNumNodes{} - 1$ nodes than,
%e.g., one or two.

%\subsubsection*{Acknowledgements}

% The author would like to thank Vaishak Belle for his comments.
%This work was supported by the EPSRC Centre for Doctoral Training in Robotics
%and Autonomous Systems, funded by the UK Engineering and Physical Sciences
%Research Council (grant EP/S023208/1).

### 2020-01-29 paper/phase_transition.tex
Finishing touches

% Created by tikzDevice version 0.12.3 on 2020-01-21 13:23:35
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,144.54);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,144.54);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (433.62,144.54);
\end{scope}
\begin{scope}
\path[clip] ( 22.14, 19.53) rectangle (430.12,141.04);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 22.14, 19.53) rectangle (430.12,141.04);
\definecolor{drawColor}{RGB}{255,255,255}

@@

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.39] at ( 17.03,100.27) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 20.39, 19.53) --
	( 22.14, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 20.39,100.38) --
	( 22.14,100.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 35.42, 17.78) --
	( 35.42, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 49.04, 17.78) --
	( 49.04, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 62.67, 17.78) --
	( 62.67, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 76.29, 17.78) --
	( 76.29, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 89.91, 17.78) --
	( 89.91, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (103.53, 17.78) --
	(103.53, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (117.15, 17.78) --
	(117.15, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (130.78, 17.78) --
	(130.78, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (144.40, 17.78) --
	(144.40, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (158.02, 17.78) --
	(158.02, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (171.64, 17.78) --
	(171.64, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (185.26, 17.78) --
	(185.26, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (198.89, 17.78) --
	(198.89, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (212.51, 17.78) --
	(212.51, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (226.13, 17.78) --
	(226.13, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (239.75, 17.78) --
	(239.75, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (253.37, 17.78) --
	(253.37, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (267.00, 17.78) --
	(267.00, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (280.62, 17.78) --
	(280.62, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (294.24, 17.78) --
	(294.24, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (307.86, 17.78) --
	(307.86, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (321.48, 17.78) --
	(321.48, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (335.11, 17.78) --
	(335.11, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (348.73, 17.78) --
	(348.73, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (362.35, 17.78) --
	(362.35, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (375.97, 17.78) --
	(375.97, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (389.59, 17.78) --
	(389.59, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (403.22, 17.78) --
	(403.22, 19.53);

\path[draw=drawColor,line width= 0.4pt,line join=round] (416.84, 17.78) --
	(416.84, 19.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,144.54);
\definecolor{drawColor}{gray}{0.30}

>>>>

% Created by tikzDevice version 0.12.3 on 2020-01-25 19:15:40
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,144.54);
\begin{scope}
\path[clip] ( 22.14, 19.53) rectangle (430.12,141.04);
\definecolor{drawColor}{gray}{0.92}

@@

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.39] at ( 17.03,100.27) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,144.54);
\definecolor{drawColor}{gray}{0.30}
